PANCHANGE_descriptives

### 2020-11-10 PANCHANGE_exclusion_descriptives.Rmd
Made the correlations for prepan and retro gad scores and their scatterplots

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("broom")
#install.packages("skimr")
#install.packages("polycor")
#install.packages("corrplot")
#install.packages("patchwork")
#install.packages("ggpubr")
```

@@

#HR

```{r Accuray scores for gad: worse/better than pre-pandemic}

dat.raw$gad.pandemic.change.group <- ifelse(dat.raw$gad.pandemic_felt_feelings_numeric <3,1,
                                           ifelse(dat.raw$gad.pandemic_felt_feelings_numeric > 3,2,0))

dat.raw$gad.pandemic.change.group <- factor(dat.raw$gad.pandemic.change.group,
                                           levels = c(1,0,2),labels = c("Worse","No change","Better"))

dat.raw <- dat.raw %>%
    mutate(gad.change.accurate = case_when(dat.raw$gad.pandemic.change.group == "Better" &
                                             dat.raw$gad.diff_score_base_prepan < 0 ~1,
                                           dat.raw$gad.pandemic.change.group == "Worse" &
                                             dat.raw$gad.diff_score_base_prepan > 0 ~ 1,
                                           dat.raw$gad.pandemic.change.group == "Better" &
                                             dat.raw$gad.diff_score_base_prepan > 0 ~ 0,
                                           dat.raw$gad.pandemic.change.group == "Worse" &
                                             dat.raw$gad.diff_score_base_prepan < 0 ~ 0))

dat.raw <- dat.raw %>%
    mutate(gad.change.retro.accurate = if_else(dat.raw$gad.pandemic.change.group == "Better" &
                                             dat.raw$gad.diff_score_retro_prepan == 0 |
                                             dat.raw$gad.pandemic.change.group == "Worse" &
                                             dat.raw$gad.diff_score_retro_prepan == 0, 1, 3 ))
dat.raw <- dat.raw %>%
  mutate(gad.change.retro.accurate = if_else(dat.raw$gad.pandemic.change.group == "Better" &
                                             dat.raw$gad.diff_score_retro_prepan >= -3 &
                                             dat.raw$gad.diff_score_retro_prepan <= 3 &
                                             dat.raw$gad.change.retro.accurate_temp != 1 |
                                             dat.raw$gad.pandemic.change.group == "Worse" &
                                             dat.raw$gad.diff_score_retro_prepan >= -3 &
                                             dat.raw$gad.diff_score_retro_prepan <= 3 &
                                             dat.raw$gad.change.retro.accurate_temp != 1,  2, gad.change.retro.accurate))

#SB - Impute no change prepandemic scores

I interpreted the presence of an NA in the retrospective score as an indicator of saying "no change" in their baseline scores. Hence, if this was the case, I imputed their baseline score. However, this interpretation could be wrong.

Also, I'm not sure what to do if there is an NA for their baseline sum score (nothing to duplicate). This was the case for ocir for some reason: the retro scores were provided for some by not the corresponding baseline score?

@@

```{r Correlation prep chunk}
# Select items for correlation matrix
prepan_retro_items <- c(
    "gad.sum_score_prepan",
    "gad.sum_score_retro", 
    "ocir.sum_score_prepan",
    "ocir.sum_score_retro",
    "phq.sum_score_prepan",
    "phq.sum_score_retro"
)

#Create a data set with only the COPING samples
prepan_retro_dat <- dat[str_detect(dat$Sample, "NBR|GLAD|EDGI"), ]

#Check that RAMP has been removed
prepan_retro_dat %>% 
  freq(Sample)

@@

# Save the p values from the correlation matrix in an object
prepan_retro.p.matrix <- as.matrix(prepan_retro.corr.mat$tests)
```

```{r Correlation matrix of prepan base sum scores}
corrplot(prepan_retro.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = prepan_retro.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )
```

>>>>

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("broom")
#install.packages("skimr")
#install.packages("polycor")
#install.packages("corrplot")
#install.packages("patchwork")
#instlal.packages("ggpubr")
```

@@

#Impute no change prepandemic scores

@@

```{r function to create upper and lower CI matrices}

cor.mtest <- function(mat, conf.level = 0.95){
mat <- as.matrix(mat)
n <- ncol(mat)
p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
diag(p.mat) <- 0
diag(lowCI.mat) <- diag(uppCI.mat) <- 1
for(i in 1:(n-1)){
for(j in (i+1):n){
tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
}
}
return(list(p.mat, lowCI.mat, uppCI.mat))
}

```{r Correlation prep chunk}
# Select items for correlation matrix
prepan_retro_items <- c(
    "gad.sum_score_prepan",
    "gad.sum_score_retro", 
    "phq.sum_score_prepan",
    "phq.sum_score_retro"
)

#Create a data set with only the GLAD and EDGI
prepan_retro_dat <- dat[str_detect(dat$Sample, "GLAD|EDGI"), ]

#Check that RAMP and NBR has been removed
prepan_retro_dat %>% 
  freq(Sample)

@@

# Save the p values from the correlation matrix in an object
prepan_retro.p.matrix <- as.matrix(prepan_retro.corr.mat$tests)

res1 <- cor.ci(prepan_retro.corr.matrix)
lowci <- cor.plot.upperLowerCi(res1)
lowci[upper.tri(lowci)] = t(lowci)[upper.tri(lowci)]

upci <- cor.plot.upperLowerCi(res1)
upci[lower.tri(upci)] = t(upci)[lower.tri(lowci)]

conf <- paste0("[", format(lowci, digits=2), " - ", format(upci, digits=2), "]")

xs <- row(lowci)
ys <- (ncol(lowci)+1) - col(lowci)

xs[lower.tri(xs,diag=TRUE)] <- 0
ys[lower.tri(ys,diag=TRUE)] <- 0

```{r Correlation matrix of prepan base sum scores}
corrplot(prepan_retro.corr.matrix, 
         method = "color", # objects to represent the correlations on plot
         type = "lower", # only use the lower triangle of the matrix
         diag = FALSE, # do not show the correlations on the diagonal
         addgrid.col = NA,
         addCoef.col = "black", # colour for the correlation coefficients in the plot
         tl.cex = 0.8,
         tl.col = "black",
         col=colorRampPalette(c("dodgerblue4","white","firebrick4"))(200), # colours for correlations
         # Combine with significance
         p.mat = prepan_retro.p.matrix, # Matrix with p values
         sig.level = 0.01, # Choose significant level
         insig = "blank" # Nonsignificant correlations have no colour
         )

text(xs, ys, conf, pos=1, cex=0.8)

#Scatterplots for self-report pandemic change

```{r Scatterplots for gad self-reported change individually}

#Filter by self-reported change since pandemic started

much_worse_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much worse")

little_worse_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "A little worse")

little_better_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "A little better")

much_better_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much better")

#Visualise each as a scatterplot

much_worse_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemc and retrospective GAD scores\nfor participants reporting much worse",
                      sep=" "),
        subtitle = paste("n", dim(much_worse_gad)[1]))

little_worse_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemc and retrospective GAD scores\nfor participants reporting a little worse",
                      sep=" "),
        subtitle = paste("n", dim(little_worse_gad)[1]))

little_better_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemc and retrospective GAD scores\nfor participants reporting a little better",
                      sep=" "),
        subtitle = paste("n", dim(little_better_gad)[1]))

much_better_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemc and retrospective GAD scores\nfor participants reporting a little worse",
                      sep=" "),
        subtitle = paste("n", dim(much_better_gad)[1]))

```{r Scatterplots for gad worse and better grouped}

#Group pt reporting much worse and a little worse, and group pt reporting much better and a little better

much_worse_little_worse_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much worse" | 
           gad.pandemic_felt_feelings == "A little worse")

much_better_little_better_gad <- dat %>% 
  filter(gad.pandemic_felt_feelings == "Much better" | 
           gad.pandemic_felt_feelings == "A little better")

#Visualise the gad sum scores as scatterplots

much_worse_little_worse_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +  geom_point(colour = "blue") +
  stat_smooth(method = "lm") +
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemc and retrospective GAD scores\nfor participants reporting much worse and a little worse",
                      sep=" "),
        subtitle = paste("n", dim(much_worse_little_worse_gad)[1]))

much_better_little_better_gad %>% ggplot(aes(x = gad.sum_score_prepan, y = gad.sum_score_retro)) +
  geom_point(colour = "blue") +
  stat_smooth(method = "lm") + 
  geom_jitter(w = 0.4, h = 0.4) +
   labs(title = paste("Relationship between prepandemc and retrospective GAD scores\nfor participants reporting much better and a little better",
                      sep=" "),
        subtitle = paste("n", dim(much_better_little_better_gad)[1]))
